---
title: "IKDDS Haemodialysis Audit Dashboard Mockup"
subtitle: "Adults on in-centre haemodialysis (ICHD) in the Republic of Ireland"
format:
  html:
    theme:
      - cosmo
      - custom.scss
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    embed-resources: true
    fig-align: center
    fig-width: 16
    fig-height: 10
    fig-dpi: 96
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(dplyr)
library(tidyr)
library(tibble)
library(rlang)
library(cli)
library(readr)
library(knitr)

# Load package
pkg_root <- normalizePath(file.path(getwd(), "../.."), mustWork = FALSE)
loaded <- FALSE
if (requireNamespace("ikdds.dashboard", quietly = TRUE)) {
  library(ikdds.dashboard)
  loaded <- TRUE
}
if (!loaded && file.exists(file.path(pkg_root, "DESCRIPTION")) &&
    requireNamespace("pkgload", quietly = TRUE)) {
  loaded <- tryCatch({
    pkgload::load_all(pkg_root)
    TRUE
  }, error = function(e) FALSE)
}
if (!loaded && file.exists(file.path(pkg_root, "R"))) {
  r_files <- list.files(file.path(pkg_root, "R"), pattern = "\\.R$",
                        full.names = TRUE)
  priority <- c("theme", "config", "data-synthetic", "metrics-helpers",
                "metrics-", "plot-helpers", "plot-")
  ordered <- character()
  for (p in priority) {
    matched <- r_files[grepl(p, basename(r_files))]
    ordered <- c(ordered, matched)
  }
  remaining <- setdiff(r_files, ordered)
  remaining <- remaining[!grepl("^(mod-|app|run)", basename(remaining))]
  for (f in c(ordered, remaining)) {
    tryCatch(source(f, local = FALSE), error = function(e) {
      message("Skipping ", basename(f), ": ", e$message)
    })
  }
}

df <- generate_synthetic_data(seed = 42, n_total = 1200)

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)
```

# Study Outcomes

```{r}
#| label: study-outcomes
tibble::tibble(
  Domain = c("Demographics", "Primary Renal Disease", "Dialysis Adequacy",
             "CKD-MBD", "Anaemia", "Blood Pressure", "Vascular Access"),
  Key_Measures = c("Age, sex",
                   "PRD classification proportions",
                   "URR, session frequency and duration",
                   "Calcium, phosphate, PTH, potassium, bicarbonate",
                   "Haemoglobin, ferritin",
                   "Pre- and post-dialysis BP",
                   "Access type and definitive access"),
  Targets = c("N/A", "N/A", "URR >65%",
              "Ca 2.2-2.5 mmol/L; Phos 1.1-1.7 mmol/L; PTH 16-72 pmol/L",
              "Hb 10-12 g/dL; Ferritin >=200 ug/L",
              "<140/90 pre; <130/80 post",
              "AVF preferred where possible")
) |>
  knitr::kable(caption = "Summary of study outcomes and clinical targets")
```

# Demographics

::: {.panel-tabset}

## Patient Demographics

```{r}
#| label: demo-age-table
compute_age_distribution(df)$summary |>
  knitr::kable(digits = 1,
               caption = "Age statistics by centre for prevalent HD patients")
```

```{r}
#| label: demo-gender
compute_gender_breakdown(df) |>
  knitr::kable(digits = 1,
               caption = "Gender breakdown by centre")
```

## Distribution of Centres

```{r}
#| label: demo-counts
compute_patient_counts(df) |>
  knitr::kable(caption = "Number of prevalent HD patients by centre at end of reporting period")
```

```{r}
#| label: demo-age-bar
#| fig-cap: "Age distribution of prevalent HD patients by centre"
#| fig-height: 8
age_data <- compute_age_distribution(df)
plot_bar(age_data$bins, .data$centre_code, .data$pct, .data$age_band,
         title = "Age Distribution by Centre", y_label = "Percentage (%)")
```

```{r}
#| label: demo-gender-bar
#| fig-cap: "Gender breakdown of prevalent HD patients by centre"
#| fig-height: 8
gender_data <- compute_gender_breakdown(df)
plot_bar(gender_data, .data$centre_code, .data$pct, .data$gender,
         title = "Gender Breakdown by Centre", y_label = "Percentage (%)")
```

:::

# Primary Renal Diagnoses

```{r}
#| label: prd-national
compute_prd_national(df) |>
  knitr::kable(digits = 1,
               caption = "National PRD distribution by ERA-EDTA classification")
```

```{r}
#| label: prd-bar
#| fig-cap: "Primary renal diagnosis distribution by centre (ERA-EDTA classification)"
prd_data <- compute_prd_proportions(df)
plot_bar(prd_data, .data$centre_code, .data$pct, .data$prd_group,
         title = "PRD Distribution by Centre", position = "stack")
```

```{r}
#| label: prd-centre
compute_prd_proportions(df) |>
  tidyr::pivot_wider(
    id_cols = "prd_group",
    names_from = "centre_code",
    values_from = "pct",
    values_fill = 0
  ) |>
  knitr::kable(digits = 1,
               caption = "PRD group percentage by centre")
```

# Adequacy of dialysis in prevalent adult ICHD patients

- Measure: Percentage of patients with URR >65%

## Adequacy of dialysis among all prevalent ICHD patients

```{r}
#| label: dial-freq
compute_session_frequency(df) |>
  tidyr::pivot_wider(
    id_cols = c("centre_code", "centre_name"),
    names_from = "hd_freq",
    values_from = "pct",
    values_fill = 0
  ) |>
  knitr::kable(digits = 1,
               caption = "Distribution of HD session frequency by centre (% of patients)")
```

```{r}
#| label: dial-duration
compute_session_duration(df) |>
  tidyr::pivot_wider(
    id_cols = c("centre_code", "centre_name"),
    names_from = "duration_band",
    values_from = "pct",
    values_fill = 0
  ) |>
  knitr::kable(digits = 1,
               caption = "Distribution of HD session duration by centre (% of patients)")
```

```{r}
#| label: dial-urr-median
#| fig-cap: "Median urea reduction ratio (URR) achieved in all prevalent adult HD patients by centre"
compute_urr_median(df) |>
  plot_caterpillar_median(x_label = "URR (%)")
```

```{r}
#| label: dial-urr-achieve
#| fig-cap: "Percentage of all prevalent HD patients with URR >65% by centre"
compute_urr_achievement(df) |>
  plot_caterpillar_proportion(target = 0.65)
```

```{r}
#| label: dial-urr-funnel
#| fig-cap: "Percentage of all prevalent HD patients with URR >65% by centre"
compute_urr_achievement(df) |> plot_funnel()
```

## Adequacy of dialysis among prevalent ICHD patients dialysing three times per week

```{r}
#| label: dial-urr-median-3x
#| fig-cap: "Median URR achieved in prevalent adult HD patients dialysing 3x/week by centre"
compute_urr_median(df, filter_3x = TRUE) |>
  plot_caterpillar_median(x_label = "URR (%)")
```

```{r}
#| label: dial-urr-achieve-3x
#| fig-cap: "Percentage of prevalent HD patients dialysing 3x/week with URR >65% by centre"
compute_urr_achievement(df, filter_3x = TRUE) |>
  plot_caterpillar_proportion(target = 0.65)
```

```{r}
#| label: dial-urr-funnel-3x
#| fig-cap: "Percentage of prevalent HD patients dialysing 3x/week with URR >65% by centre"
compute_urr_achievement(df, filter_3x = TRUE) |> plot_funnel()
```

# Blood Pressure

- Measure: Proportion of patients with BP <140/90 (pre-dialysis) and/or <130/80 (post-dialysis)

```{r}
#| label: bp-summary
compute_bp_summary(df) |>
  knitr::kable(digits = 1,
               caption = "Blood pressure summary for prevalent patients on HD by centre")
```

```{r}
#| label: bp-caterpillar
#| fig-cap: "Percentage of adult patients prevalent to ICHD with BP <140/90 (pre-HD) and/or <130/80 (post-HD)"
#| fig-height: 12
pre <- compute_pre_bp_achievement(df)
post <- compute_post_bp_achievement(df)

p1 <- plot_caterpillar_proportion(pre, title = "Pre-HD BP <140/90")
p2 <- plot_caterpillar_proportion(post, title = "Post-HD BP <130/80")

if (requireNamespace("patchwork", quietly = TRUE)) {
  patchwork::wrap_plots(p1, p2, ncol = 1)
} else {
  print(p1)
  print(p2)
}
```

```{r}
#| label: bp-funnel-pre
#| fig-cap: "Percentage of adult patients prevalent to ICHD with BP <140/90 (pre-HD) by centre"
compute_pre_bp_achievement(df) |> plot_funnel()
```

```{r}
#| label: bp-funnel-post
#| fig-cap: "Percentage of adult patients prevalent to ICHD with BP <130/80 (post-HD) by centre"
compute_post_bp_achievement(df) |> plot_funnel()
```

# Biochemistry parameters in prevalent adult ICHD patients

## Serum Phosphate

```{r}
#| label: bio-po4-caterpillar
#| fig-cap: "Percentage of adult patients prevalent to ICHD with Serum Phosphate 1.1-1.7 mmol/L by centre. CI - confidence interval"
compute_phosphate_achievement(df) |> plot_caterpillar_proportion()
```

```{r}
#| label: bio-po4-funnel
#| fig-cap: "Percentage of adult patients prevalent to ICHD with Serum Phosphate 1.1-1.7 mmol/L by centre"
compute_phosphate_achievement(df) |> plot_funnel()
```

## Adjusted Calcium

```{r}
#| label: bio-ca-caterpillar
#| fig-cap: "Percentage of adult patients prevalent to ICHD with adjusted calcium (Ca) 2.2-2.5 mmol/L by centre. CI - confidence interval"
compute_calcium_achievement(df) |> plot_caterpillar_proportion()
```

```{r}
#| label: bio-ca-funnel
#| fig-cap: "Percentage of adult patients prevalent to ICHD with adjusted calcium (Ca) 2.2-2.5 mmol/L by centre"
compute_calcium_achievement(df) |> plot_funnel()
```

## Parathyroid Hormone (PTH)

```{r}
#| label: bio-pth-caterpillar
#| fig-cap: "Percentage of adult patients prevalent to ICHD with PTH 16-72 pmol/L by centre. CI - confidence interval"
compute_pth_achievement(df) |> plot_caterpillar_proportion()
```

```{r}
#| label: bio-pth-funnel
#| fig-cap: "Percentage of adult patients prevalent to ICHD with PTH 16-72 pmol/L by centre"
compute_pth_achievement(df) |> plot_funnel()
```

## Simultaneous control of adjusted calcium, phosphate and PTH

```{r}
#| label: bio-mbd-caterpillar
#| fig-cap: "Percentage of adult patients prevalent to ICHD achieving simultaneous control of all three CKD-MBD targets"
compute_ckd_mbd_simultaneous(df) |> plot_caterpillar_proportion()
```

# Pre-dialysis Bicarbonate and Potassium in prevalent adult HD patients

```{r}
#| label: bk-k-caterpillar
#| fig-cap: "Percentage of adult patients prevalent to ICHD with pre-dialysis potassium (K) 4-6 mmol/L by centre. CI - confidence interval"
compute_potassium_achievement(df) |> plot_caterpillar_proportion()
```

```{r}
#| label: bk-k-funnel
#| fig-cap: "Percentage of adult patients prevalent to ICHD with pre-dialysis potassium (K) 4-6 mmol/L by centre"
compute_potassium_achievement(df) |> plot_funnel()
```

```{r}
#| label: bk-hco3-caterpillar
#| fig-cap: "Percentage of adult patients prevalent to ICHD with pre-dialysis Bicarbonate 18-26 mmol/L by centre. CI - confidence interval"
compute_bicarbonate_achievement(df) |> plot_caterpillar_proportion()
```

```{r}
#| label: bk-hco3-funnel
#| fig-cap: "Percentage of adult patients prevalent to ICHD with pre-dialysis Bicarbonate 18-26 mmol/L by centre"
compute_bicarbonate_achievement(df) |> plot_funnel()
```

# Anaemia in prevalent adult ICHD patients

```{r}
#| label: ana-hb-median
#| fig-cap: "Median haemoglobin (Hb) in adult patients prevalent to ICHD by centre"
compute_hb_median(df) |>
  plot_caterpillar_median(x_label = "Hb (g/dL)")
```

```{r}
#| label: ana-ferritin-caterpillar
#| fig-cap: "Percentage of adult patients prevalent to ICHD with ferritin >=200 ug/L by centre. CI - confidence interval"
compute_ferritin_achievement(df) |> plot_caterpillar_proportion()
```

```{r}
#| label: ana-ferritin-funnel
#| fig-cap: "Percentage of adult patients prevalent to ICHD with ferritin >=200 ug/L by centre"
compute_ferritin_achievement(df) |> plot_funnel()
```

```{r}
#| label: ana-hb-dist
#| fig-cap: "Distribution of haemoglobin (Hb) in adult patients prevalent to ICHD by centre"
plot_distribution(df, .data$qble1,
                  title = "Haemoglobin Distribution",
                  x_label = "Hb (g/dL)",
                  target_lower = 10, target_upper = 12)
```

```{r}
#| label: ana-hb-achieve
#| fig-cap: "Percentage of adult patients prevalent to ICHD with Haemoglobin 10-12 g/dL by centre"
compute_hb_achievement(df) |> plot_caterpillar_proportion()
```

```{r}
#| label: ana-hb-funnel
#| fig-cap: "Percentage of adult patients prevalent to ICHD with Haemoglobin 10-12 g/dL by centre"
compute_hb_achievement(df) |> plot_funnel()
```

# Dialysis access in prevalent adult dialysis patients

```{r}
#| label: acc-bar
#| fig-cap: "Vascular access type distribution by centre"
access_data <- compute_access_distribution(df)
plot_bar(access_data, .data$centre_code, .data$pct, .data$access_type,
         title = "Vascular Access Type", position = "stack")
```

```{r}
#| label: acc-avf
#| fig-cap: "AVF rate by centre. CI - confidence interval"
compute_avf_rate(df) |> plot_caterpillar_proportion()
```

```{r}
#| label: acc-table
compute_access_distribution(df) |>
  tidyr::pivot_wider(
    id_cols = c("centre_code", "centre_name"),
    names_from = "access_type",
    values_from = c("count", "pct"),
    values_fill = 0
  ) |>
  knitr::kable(digits = 1,
               caption = "Vascular access type by centre")
```

# Appendix 1 {.appendix}

## Missing Data

```{r}
#| label: missing
#| fig-cap: "Missing Data by Centre"
compute_missing_national(df) |>
  knitr::kable(digits = 1,
               caption = "Missing data rates for key variables")
```

# Appendix 2 {.appendix}

## Caterpillar plots

Caterpillar plots show the **centre-level estimate** (dot) with its **95% confidence interval** (vertical whisker). Centres are ordered from highest to lowest (left to right). The **solid red line** shows the national weighted average. Where applicable, a **dotted green line** shows the clinical target.

Confidence intervals for **proportions** use the Wilson score method. Confidence intervals for **medians** use the order-statistic method.

## Funnel plots

Funnel plots (Spiegelhalter method) show **centre proportion** (y-axis) against **centre size** (x-axis). The **solid orange line** is the national average. The **solid curves** are the 95% control limits; the **dashed curves** are the 99.7% control limits. Centres outside these limits show statistically unusual variation.

# Appendix 3 {.appendix}

## Patient-Level Data

```{r}
#| label: drilldown
df |>
  dplyr::select(
    ID = record_id, Centre = centre_code, Consultant = consultant,
    Age = age, Gender = gender, `URR (%)` = qblg9,
    `Pre SBP` = qblg3, `Pre DBP` = qblg4,
    `Post SBP` = qblg6, `Post DBP` = qblg7,
    PO4 = qblb1, Ca = qblb4, PTH = qblb9,
    `K+` = qbla9, HCO3 = qbla4, Hb = qble1,
    Ferritin = qblf1, Access = qhd20
  ) |>
  head(50) |>
  knitr::kable(digits = 1,
               caption = "Patient-level data (first 50 rows)")
```
