---
title: "IKDDS Dashboard Specification"
subtitle: "Technical specification for Power BI recreation by HSE IIS"
author: "IKDDS Team"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    theme: cosmo
    code-fold: true
    code-tools: true
    embed-resources: true
params:
  registry_path: null
  data_path: null
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(dplyr)
library(tidyr)
library(tibble)
library(rlang)
library(cli)
library(readr)
library(knitr)

# Load package: try library() first, then pkgload, then source R/ files
pkg_root <- normalizePath(file.path(getwd(), "../.."), mustWork = FALSE)
loaded <- FALSE
if (requireNamespace("ikdds.dashboard", quietly = TRUE)) {
  library(ikdds.dashboard)
  loaded <- TRUE
}
if (!loaded && file.exists(file.path(pkg_root, "DESCRIPTION")) &&
    requireNamespace("pkgload", quietly = TRUE)) {
  loaded <- tryCatch({
    pkgload::load_all(pkg_root)
    TRUE
  }, error = function(e) FALSE)
}
if (!loaded && file.exists(file.path(pkg_root, "R"))) {
  r_files <- list.files(file.path(pkg_root, "R"), pattern = "\\.R$",
                        full.names = TRUE)
  priority <- c("theme", "config", "data-synthetic", "metrics-helpers",
                "metrics-", "plot-helpers", "plot-")
  ordered <- character()
  for (p in priority) {
    matched <- r_files[grepl(p, basename(r_files))]
    ordered <- c(ordered, matched)
  }
  remaining <- setdiff(r_files, ordered)
  remaining <- remaining[!grepl("^(mod-|app|run)", basename(remaining))]
  for (f in c(ordered, remaining)) {
    tryCatch(source(f, local = FALSE), error = function(e) {
      message("Skipping ", basename(f), ": ", e$message)
    })
  }
}

# Load registry and data from params or generate fresh
if (!is.null(params$registry_path) && file.exists(params$registry_path)) {
  registry <- readRDS(params$registry_path)
} else {
  registry <- dashboard_registry()
}

if (!is.null(params$data_path) && file.exists(params$data_path)) {
  df <- readRDS(params$data_path)
} else {
  df <- generate_synthetic_data(seed = 42, n_total = 1200)
}
```

## Introduction

This document is the **technical specification** for the IKDDS Haemodialysis Audit Dashboard. It is auto-generated from the dashboard registry — the single source of truth that defines every view, card, metric, and variable mapping.

**Purpose:** Provide HSE Integrated Information Services (IIS) with all information needed to recreate this dashboard in Power BI, including:

- Variable mappings (REDCap field → eMed source → display label)
- Metric formulas in both R code and pseudocode (for DAX/Power Query translation)
- Chart types and example figures
- Statistical methods and formulae
- Filter specifications
- HSE colour palette

## Colour Palette

```{r}
#| label: colours
registry_colours() |> knitr::kable()
```

## Statistical Methods

```{r}
#| label: methods
registry_stat_methods(registry) |> knitr::kable()
```

## Global Filters

The following filters apply across all dashboard views:

```{r}
#| label: filters
registry_filters(registry) |> knitr::kable()
```

## Variable Dictionary (All Views)

Complete mapping of REDCap fields to eMed sources:

```{r}
#| label: all-variables
registry_fields(registry) |>
  dplyr::select(view_id, card_id, redcap, label, emed_table, emed_col, unit,
                target_lower, target_upper) |>
  dplyr::distinct(redcap, .keep_all = TRUE) |>
  knitr::kable()
```

---

```{r}
#| label: view-sections
#| output: asis
#| fig-width: 8
#| fig-height: 5

for (view in registry) {
  cat("\n\n## ", view$view_title, "\n\n")
  cat("**Description:** ", view$description, "\n\n")
  cat("**Filters:** ", paste(view$filters, collapse = ", "), "\n\n")

  # Variable dictionary for this view
  view_fields <- list()
  for (card in view$cards) {
    for (field in card$fields) {
      view_fields <- c(view_fields, list(tibble::tibble_row(
        REDCap = field$redcap,
        Label  = field$label,
        `eMed Table` = field$emed_table,
        `eMed Column` = field$emed_col,
        Unit   = field$unit %||% "",
        `Target Lower` = ifelse(is.na(field$target_lower), "", as.character(field$target_lower)),
        `Target Upper` = ifelse(is.na(field$target_upper), "", as.character(field$target_upper))
      )))
    }
  }

  if (length(view_fields) > 0) {
    fields_tbl <- dplyr::bind_rows(view_fields) |> dplyr::distinct()
    cat("\n### Variable Dictionary\n\n")
    print(knitr::kable(fields_tbl))
    cat("\n")
  }

  # Per-card sections
  for (card in view$cards) {
    cat("\n### ", card$title, "\n\n")
    cat("- **Card ID:** `", card$card_id, "`\n")
    cat("- **Card Type:** ", card$card_type, "\n")
    if (!is.null(card$chart_type)) {
      cat("- **Chart Type:** ", card$chart_type, "\n")
    }
    if (!is.null(card$metric_fn)) {
      cat("- **Metric Function:** `", card$metric_fn, "()`\n")
    }
    if (!is.null(card$plot_fn)) {
      cat("- **Plot Function:** `", card$plot_fn, "()`\n")
    }
    if (!is.null(card$stat_method)) {
      cat("- **Statistical Method:** ", card$stat_method, "\n")
    }
    if (!is.null(card$target)) {
      cat("- **Target:** ", card$target, "\n")
    }
    cat("\n")

    # R Code
    if (!is.null(card$r_code) && !is.na(card$r_code)) {
      cat("\n**R Code:**\n\n")
      cat("```r\n")
      cat(card$r_code, "\n")
      cat("```\n\n")
    }

    # Pseudocode
    if (!is.null(card$pseudocode) && !is.na(card$pseudocode)) {
      cat("\n**Pseudocode (for DAX/Power Query):**\n\n")
      cat("> ", card$pseudocode, "\n\n")
    }

    # Example figure
    if (!is.null(card$plot_fn) && card$card_type == "chart") {
      cat("\n**Example Figure:**\n\n")

      tryCatch({
        if (card$chart_type == "bar") {
          if (card$card_id == "demo_age_bar") {
            p <- plot_bar(compute_age_distribution(df)$bins,
                          .data$centre_code, .data$pct, .data$age_band,
                          title = card$title, y_label = "Percentage (%)")
          } else if (card$card_id == "demo_gender_bar") {
            p <- plot_bar(compute_gender_breakdown(df),
                          .data$centre_code, .data$pct, .data$gender,
                          title = card$title, y_label = "Percentage (%)")
          } else if (card$card_id == "prd_bar") {
            p <- plot_bar(compute_prd_proportions(df),
                          .data$centre_code, .data$pct, .data$prd_group,
                          title = card$title, position = "stack")
          } else if (card$card_id == "acc_bar") {
            p <- plot_bar(compute_access_distribution(df),
                          .data$centre_code, .data$pct, .data$access_type,
                          title = card$title, position = "stack")
          }
          print(p)
        } else if (!is.null(card$metric_fn)) {
          metric_fn <- get(card$metric_fn)
          plot_fn <- get(card$plot_fn)
          metric_data <- metric_fn(df)

          if (card$chart_type == "caterpillar_proportion") {
            if (!is.null(card$target)) {
              p <- plot_fn(metric_data, target = card$target)
            } else {
              p <- plot_fn(metric_data)
            }
          } else if (card$chart_type == "caterpillar_median") {
            x_lab <- if (card$card_id == "dial_urr_median") "URR (%)"
                     else if (card$card_id == "ana_hb_median") "Hb (g/dL)"
                     else "Value"
            p <- plot_fn(metric_data, x_label = x_lab)
          } else if (card$chart_type == "funnel") {
            p <- plot_fn(metric_data)
          } else {
            p <- plot_fn(metric_data)
          }
          print(p)
        } else if (card$chart_type == "distribution") {
          p <- plot_distribution(df, .data$qble1,
                                 title = card$title,
                                 x_label = "Hb (g/dL)",
                                 target_lower = 10, target_upper = 12)
          print(p)
        }
      }, error = function(e) {
        cat("\n*Figure generation failed:* ", conditionMessage(e), "\n\n")
      })
    }
  }
}
```
