---
title: "IKDDS Dashboard Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{IKDDS Dashboard Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The `ikdds.dashboard` package provides an interactive Shiny dashboard for
the Irish Kidney Disease Data System (IKDDS) annual haemodialysis audit.
It provides **14 dashboard pages** rendering 23 clinical metrics across
7 domains as caterpillar plots, funnel plots, centre profiles, individual
patient QA comparisons, and patient-level drill-down tables.

## Quick Start

```{r eval=FALSE}
library(ikdds.dashboard)

# Launch with synthetic demo data (default)
run_app()
```

## Data Sources

The dashboard supports two data sources, controlled by the
`IKDDS_DASH_DATA_SOURCE` environment variable:

- **synthetic** (default): Generates ~1,200 realistic patients across
  7 Irish renal centres. No credentials needed.
- **redcap**: Reads live data from the IKDDS REDCap project.
  Requires `IKDDS_REDCAP_URI` and `IKDDS_REDCAP_TOKEN`.

## Dashboard Pages

| # | Page | Description |
|---|------|-------------|
| 1 | Overview | National summary KPIs, achievement rates, missing data overview |
| 2 | Demographics | Age/gender distributions by centre |
| 3 | PRD | Primary renal diagnosis (ERA classification) |
| 4 | Dialysis | URR caterpillar + funnel, session frequency/duration |
| 5 | Blood Pressure | Pre/post-HD BP caterpillar + funnel |
| 6 | Biochemistry | PO4, Ca, PTH caterpillar + funnel |
| 7 | Bicarb / K+ | Bicarbonate and potassium caterpillar + funnel |
| 8 | Anaemia | Hb and ferritin caterpillar + funnel + distribution |
| 9 | Access | Vascular access type distribution |
| 10 | Data Quality | Completeness heatmap, flagged centres, national KPIs |
| 11 | Centre Profile | Single-centre deep-dive vs national averages |
| 12 | Individual | Patient-level QA: compare patient vs centre/national medians |
| 13 | Methods | Auto-generated metric definitions and statistical methods |
| 14 | Drill-down | Patient-level data table |

## Clinical Domains

| Domain | Metrics | Key Plots |
|--------|---------|-----------|
| Demographics | Patient counts, age, gender | Tables, bar charts |
| PRD | ERA classification | Tables, stacked bars |
| Dialysis | URR, session freq/duration | Caterpillar, funnel |
| Blood Pressure | Pre/post-HD BP | Caterpillar, funnel |
| Biochemistry | PO4, Ca, PTH, CKD-MBD | Caterpillar, funnel |
| Bicarb/K+ | Bicarbonate, potassium | Caterpillar, funnel |
| Anaemia | Hb, ferritin | Caterpillar, funnel, histogram |
| Vascular Access | AVF, AVG, catheter | Tables, bar charts |

## Key Features

- **Metrics Registry** — YAML-based centralised definitions
  (`inst/metrics/metrics.yml`) with targets, suppression thresholds,
  and compute function bindings.
- **Suppression Engine** — Automatic suppression of centres with n < 10
  patients, with visual indicators in plots and tables.
- **Statistical Methods** — Wilson score confidence intervals for
  proportions, order-statistic CIs for medians, Spiegelhalter funnel
  plots with 95% and 99.7% control limits.
- **Individual Patient QA** — Compare a single patient's lab results
  against centre and national medians, with RAG status indicators
  and bullet charts.
- **Centre Profile** — Deep-dive into a single centre's performance
  including consultant-level variation tables.

## Using Metric Functions Standalone

All metric functions are pure (tibble in, tibble out) and can be used
outside the dashboard:

```{r eval=FALSE}
config <- dashboard_config()
df <- get_audit_data(config)

# Compute URR achievement by centre
urr <- compute_urr_achievement(df)

# Create caterpillar plot
plot_caterpillar_proportion(urr, target = 0.65)

# Compute centre-level missing data rates
missing <- compute_missing_data(df)
```

## Further Reading

- `vignette("metric-definitions")` — All 23 metric definitions
- `vignette("synthetic-data")` — Generating and exploring test data
- `vignette("statistical-methods")` — CIs, funnels, and suppression
- `vignette("individual-patient-qa")` — Individual patient QA module
- `vignette("data-quality-workflow")` — Data completeness analysis
- `vignette("deployment-guide")` — Deployment options
