---
title: "IKDDS Dashboard Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{IKDDS Dashboard Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The `ikdds.dashboard` package provides an interactive Shiny dashboard for
the Irish Kidney Disease Data System (IKDDS) annual haemodialysis audit.
It renders 23 clinical metrics across 7 domains as caterpillar plots,
funnel plots, and patient-level drill-down tables.

## Quick Start

```{r eval=FALSE}
library(ikdds.dashboard)

# Launch with synthetic demo data (default)
run_app()
```

## Data Sources

The dashboard supports two data sources, controlled by the
`IKDDS_DASH_DATA_SOURCE` environment variable:

- **synthetic** (default): Generates ~1,200 realistic patients across
  7 Irish renal centres. No credentials needed.
- **redcap**: Reads live data from the IKDDS REDCap project.
  Requires `IKDDS_REDCAP_URI` and `IKDDS_REDCAP_TOKEN`.

## Clinical Domains

| Domain | Metrics | Key Plots |
|--------|---------|-----------|
| Demographics | Patient counts, age, gender | Tables, bar charts |
| PRD | ERA classification | Tables, stacked bars |
| Dialysis | URR, session freq/duration | Caterpillar, funnel |
| Blood Pressure | Pre/post-HD BP | Caterpillar, funnel |
| Biochemistry | PO4, Ca, PTH, CKD-MBD | Caterpillar, funnel |
| Bicarb/K+ | Bicarbonate, potassium | Caterpillar, funnel |
| Anaemia | Hb, ferritin | Caterpillar, funnel, histogram |
| Vascular Access | AVF, AVG, catheter | Tables, bar charts |

## Using Metric Functions Standalone

All metric functions are pure (tibble in, tibble out) and can be used
outside the dashboard:

```{r eval=FALSE}
config <- dashboard_config()
df <- get_audit_data(config)

# Compute URR achievement by centre
urr <- compute_urr_achievement(df)

# Create caterpillar plot
plot_caterpillar_proportion(urr, target = 0.65)
```
