---
title: "Data Quality Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Quality Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Overview

Data quality is critical for a reliable clinical audit. The IKDDS
dashboard includes dedicated tools for assessing completeness,
identifying problematic centres, and understanding missing data
patterns. This vignette demonstrates the data quality workflow
using the package's metric functions.

```{r}
library(ikdds.dashboard)

df <- generate_synthetic_data(seed = 42, n_total = 1200)
```

## Missing Data Analysis

### Centre-Level Missing Data

The `compute_missing_data()` function computes the percentage of
missing values for each clinical variable at each centre:

```{r}
missing <- compute_missing_data(df)
head(missing)
```

### National-Level Summary

For a national overview of missing data rates:

```{r}
missing_national <- compute_missing_national(df)
missing_national
```

## Key Clinical Variables

The dashboard tracks completeness for 13 key variables:

| Variable | Label | Clinical Domain |
|----------|-------|-----------------|
| `qblg9` | URR | Dialysis adequacy |
| `qblg3` | Pre SBP | Blood pressure |
| `qblg4` | Pre DBP | Blood pressure |
| `qblg6` | Post SBP | Blood pressure |
| `qblg7` | Post DBP | Blood pressure |
| `qblb1` | Phosphate | Biochemistry |
| `qblb4` | Adj. Calcium | Biochemistry |
| `qblb9` | PTH | Biochemistry |
| `qbla9` | Potassium | Bicarbonate/K+ |
| `qbla4` | Bicarbonate | Bicarbonate/K+ |
| `qble1` | Haemoglobin | Anaemia |
| `qblf1` | Ferritin | Anaemia |
| `qhd20` | Vascular Access | Access |

## Completeness Thresholds

The dashboard applies a 70% completeness threshold. Centres below
this threshold receive a warning label in caterpillar and funnel
plots:

```{r}
urr <- compute_urr_achievement(df)
labelled <- add_completeness_labels(urr, df, "qblg9")
labelled[, c("centre_code", "n", "label")]
```

## Suppression for Small Numbers

Centres with fewer than 10 patients are suppressed to prevent
identification risk:

```{r}
# Demonstrate with a subset
small_subset <- df[df$centre_code %in% c("BEA"), ][1:8, ]
small_subset$centre_code <- "TINY"  # Simulate a small centre
test_df <- rbind(df, small_subset)

urr <- compute_urr_achievement(test_df)
suppressed <- apply_suppression(urr, min_n = 10)
suppressed[suppressed$centre_code == "TINY",
           c("centre_code", "n", "proportion", "suppressed")]
```

## Unmapped Patients

The dashboard identifies patients without consultant mapping,
which can indicate data entry issues:

```{r}
unmapped <- compute_unmapped_patients(df)
unmapped
```

## Validation Pipeline

The `validate_machine_fields()` function checks for machine-populated
fields that may contain invalid zero values and replaces them with NA:

```{r}
# Check how many zeros exist in lab fields before validation
lab_fields <- c("qblg9", "qblb1", "qblb4", "qblb9", "qble1", "qblf1")
zeros_before <- sapply(lab_fields, function(v) sum(df[[v]] == 0, na.rm = TRUE))
zeros_before
```

```{r}
# Validate (replaces 0s with NA for machine-populated fields)
df_validated <- validate_machine_fields(df)
zeros_after <- sapply(lab_fields, function(v) {
  sum(df_validated[[v]] == 0, na.rm = TRUE)
})
zeros_after
```

## Data Quality Dashboard Page

The Data Quality page (Page 10) in the dashboard provides:

1. **National KPIs** — Overall completeness percentages
2. **Completeness heatmap** — Centre x Variable matrix showing
   missing data rates with colour coding
3. **Flagged centres** — Centres with any metric >30% missing
4. **Unmapped patients** — Patients without consultant assignment

## Integrating with Individual Patient QA

The Individual Patient QA page (Page 12) shows "Missing" status
in grey for any metric where the patient has no recorded value.
This helps identify data gaps at the individual level:

```{r}
# Check a patient's completeness
patient <- df[1, ]
metrics <- c("qblg9", "qblg3", "qblg4", "qblg6", "qblg7",
             "qblb1", "qblb4", "qblb9", "qbla9", "qbla4", "qble1")
patient_completeness <- sapply(metrics, function(v) !is.na(patient[[v]]))
data.frame(
  Metric = metrics,
  Has_Value = patient_completeness
)
```
