---
title: "Deployment Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Deployment Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Local Development

### Prerequisites

- R >= 4.1.0
- RStudio (recommended)
- Required packages: `shiny`, `bslib`, `ggplot2`, `plotly`, `DT`, `dplyr`,
  `tidyr`, `tibble`, `rlang`, `glue`, `cli`, `purrr`, `stringr`, `readr`,
  `ggrepel`, `yaml`

### Install from Source

```{r eval=FALSE}
devtools::install("ikdds_dashboard/ikdds.dashboard")
```

### Run Locally

```{r eval=FALSE}
library(ikdds.dashboard)
run_app()
```

The dashboard launches with 14 pages covering all clinical domains plus
data quality, centre profile, individual patient QA, methods, and
drill-down views.

## Environment Variables

| Variable | Default | Description |
|----------|---------|-------------|
| `IKDDS_DASH_DATA_SOURCE` | `synthetic` | `synthetic` or `redcap` |
| `IKDDS_REDCAP_URI` | — | REDCap API endpoint |
| `IKDDS_REDCAP_TOKEN` | — | REDCap API token |
| `IKDDS_DASH_CACHE_TTL` | `3600` | Cache TTL in seconds |
| `IKDDS_DASH_TITLE` | `IKDDS Haemodialysis Audit Dashboard` | App title |
| `IKDDS_DASH_DEBUG` | `false` | Enable debug mode |
| `PORT` | `3838` | HTTP port |

## Posit Connect Deployment

1. Set environment variables in the Posit Connect admin panel
2. Push via git or use `rsconnect::deployApp()`
3. The `app.R` file at the package root serves as the entrypoint

```{r eval=FALSE}
rsconnect::deployApp(
  appDir = "ikdds_dashboard/ikdds.dashboard",
  appTitle = "IKDDS Haemodialysis Audit Dashboard"
)
```

## Docker (Optional)

```dockerfile
FROM rocker/shiny:4.4.0
RUN install2.r --error shiny bslib ggplot2 plotly DT dplyr tidyr \
    tibble rlang glue cli purrr stringr readr ggrepel yaml
COPY ikdds.dashboard /srv/shiny-server/ikdds.dashboard
EXPOSE 3838
```

## Security Notes

- REDCap tokens should be stored as environment variables, never in code
- The dashboard does not write any data — it is read-only
- Patient identifiers (`record_id`) are synthetic IDs, not real MRNs
- The Individual Patient QA page displays individual patient values for
  clinical QA purposes — ensure appropriate access controls in production
- In production, configure HTTPS via reverse proxy
