---
title: "Individual Patient QA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Individual Patient QA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Purpose

The Individual Patient QA page enables nephrologists to compare a
single patient's results against centre and national averages. This
supports individual-level quality assurance to identify patients
needing clinical attention, ultimately improving centre-level
performance.

## Workflow

1. Select a **centre** from the dropdown
2. Select a **patient** (record ID) — the list cascades from the
   selected centre
3. Click **Load Patient** to generate the comparison

The page then displays:

- **KPI boxes** — Patient ID, Centre, and Consultant
- **Comparison table** — 11 metrics with patient value, centre median,
  national median, target range, and RAG status
- **Bullet chart** — Visual comparison of patient values against
  centre and national medians

## The 11 Metrics

| Metric | REDCap Field | Target | Type |
|--------|-------------|--------|------|
| URR (%) | `qblg9` | >65 | Greater than |
| Pre SBP | `qblg3` | <140 | Less than |
| Pre DBP | `qblg4` | <90 | Less than |
| Post SBP | `qblg6` | <130 | Less than |
| Post DBP | `qblg7` | <80 | Less than |
| PO4 (mmol/L) | `qblb1` | 1.1-1.7 | Range |
| Ca (mmol/L) | `qblb4` | 2.2-2.5 | Range |
| PTH (pmol/L) | `qblb9` | 16-72 | Range |
| K+ (mmol/L) | `qbla9` | 4-6 | Range |
| HCO3 (mmol/L) | `qbla4` | 18-26 | Range |
| Hb (g/dL) | `qble1` | 10-12 | Range |

## RAG Status Logic

Each metric is classified using a simple target-checking function:

- **In range** (green) — patient value meets the target criterion
- **Out of range** (red, bold) — patient value is outside the target
- **Missing** (grey) — no value recorded for this patient

```{r}
library(ikdds.dashboard)

df <- generate_synthetic_data(seed = 42, n_total = 1200)

# Pick a patient from Beaumont
bea_patients <- df[df$centre_code == "BEA", ]
patient <- bea_patients[1, ]

cat("Patient:", patient$record_id, "\n")
cat("Centre:", patient$centre_code, "\n")
cat("Consultant:", patient$consultant, "\n")
```

## Computing Comparisons

The module uses simple `median()` calls — no dependency on the
metrics engine:

```{r}
# Centre medians (Beaumont)
centre_df <- df[df$centre_code == "BEA", ]

# National medians (all centres)
national_df <- df

# Build comparison for one metric: URR
patient_urr <- patient$qblg9
centre_urr <- median(centre_df$qblg9, na.rm = TRUE)
national_urr <- median(national_df$qblg9, na.rm = TRUE)

cat("Patient URR:", round(patient_urr, 1), "\n")
cat("Centre median URR:", round(centre_urr, 1), "\n")
cat("National median URR:", round(national_urr, 1), "\n")
cat("Target: >65\n")
cat("Status:", ifelse(is.na(patient_urr), "Missing",
                       ifelse(patient_urr > 65, "In range", "Out of range")), "\n")
```

## Building the Full Comparison Table

Here is how the module builds the comparison for all 11 metrics:

```{r}
metric_defs <- list(
  list(label = "URR (%)",      col = "qblg9", lo = 65,   hi = Inf,  op = "gt"),
  list(label = "Pre SBP",      col = "qblg3", lo = -Inf, hi = 140,  op = "lt"),
  list(label = "Pre DBP",      col = "qblg4", lo = -Inf, hi = 90,   op = "lt"),
  list(label = "Post SBP",     col = "qblg6", lo = -Inf, hi = 130,  op = "lt"),
  list(label = "Post DBP",     col = "qblg7", lo = -Inf, hi = 80,   op = "lt"),
  list(label = "PO4 (mmol/L)", col = "qblb1", lo = 1.1,  hi = 1.7,  op = "range"),
  list(label = "Ca (mmol/L)",  col = "qblb4", lo = 2.2,  hi = 2.5,  op = "range"),
  list(label = "PTH (pmol/L)", col = "qblb9", lo = 16,   hi = 72,   op = "range"),
  list(label = "K+ (mmol/L)",  col = "qbla9", lo = 4,    hi = 6,    op = "range"),
  list(label = "HCO3 (mmol/L)",col = "qbla4", lo = 18,   hi = 26,   op = "range"),
  list(label = "Hb (g/dL)",    col = "qble1", lo = 10,   hi = 12,   op = "range")
)

in_target <- function(val, m) {
  if (is.na(val)) return(NA)
  switch(m$op,
    gt    = val > m$lo,
    lt    = val < m$hi,
    range = val >= m$lo & val <= m$hi
  )
}

format_target <- function(m) {
  switch(m$op,
    gt    = paste0(">", m$lo),
    lt    = paste0("<", m$hi),
    range = paste0(m$lo, "-", m$hi)
  )
}

comparison <- do.call(rbind, lapply(metric_defs, function(m) {
  pt_val <- patient[[m$col]]
  centre_med <- median(centre_df[[m$col]], na.rm = TRUE)
  national_med <- median(national_df[[m$col]], na.rm = TRUE)

  status <- if (is.na(pt_val)) "Missing"
            else if (in_target(pt_val, m)) "In range"
            else "Out of range"

  data.frame(
    Metric = m$label,
    Patient = round(pt_val, 1),
    Centre_Median = round(centre_med, 1),
    National_Median = round(national_med, 1),
    Target = format_target(m),
    Status = status,
    stringsAsFactors = FALSE
  )
}))

comparison
```

## Bullet Chart

The bullet chart provides a visual comparison. For each metric with
a non-missing patient value:

- **Grey bar** (wide) — national median
- **Teal bar** (narrow) — patient value
- **Orange marker** — centre median
- **Red marker** — national median

This follows the same visual pattern as the Centre Profile bullet
chart, adapted for individual patient data.

## Clinical Use Case

A nephrologist reviewing their centre's audit results can:

1. Go to the **Individual** tab
2. Select their centre and a specific patient
3. Immediately see which metrics are out of range (red)
4. Identify whether the patient is above or below centre norms
5. Prioritise clinical interventions for that patient

When aggregated across patients, this workflow helps identify
systematic issues contributing to centre-level underperformance
on specific metrics.
