---
title: "Synthetic Data Generation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Synthetic Data Generation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

The `ikdds.dashboard` package includes a synthetic data generator that
creates realistic haemodialysis patient data for development, testing,
and demonstration. This vignette shows how to generate data, inspect
its structure, and use it with the metric functions.

## Generating Synthetic Data

```{r}
library(ikdds.dashboard)

# Generate ~1,200 patients across 7 Irish renal centres
df <- generate_synthetic_data(seed = 42, n_total = 1200)

cat("Rows:", nrow(df), "\n")
cat("Columns:", ncol(df), "\n")
```

## Data Structure

Each row represents one patient in one audit period. Key columns:

```{r}
str(df[, c("record_id", "centre_code", "centre_name", "consultant",
           "age", "gender")])
```

### Clinical Variables

The dataset contains the REDCap field codes used by the IKDDS audit:

```{r}
# Dialysis adequacy
summary(df$qblg9)  # URR (%)

# Blood pressure
summary(df[, c("qblg3", "qblg4", "qblg6", "qblg7")])
```

```{r}
# Biochemistry
summary(df[, c("qblb1", "qblb4", "qblb9")])  # PO4, Ca, PTH
```

```{r}
# Anaemia
summary(df[, c("qble1", "qblf1")])  # Hb, Ferritin
```

## Centre Distribution

```{r}
table(df$centre_code)
```

The 7 centres reflect the real Irish renal network:

- **BEA** — Beaumont Hospital (Dublin, largest centre)
- **MAT** — Mater Misericordiae University Hospital (Dublin)
- **TAL** — Tallaght University Hospital (Dublin)
- **STJ** — St. James's Hospital (Dublin)
- **CUH** — Cork University Hospital (Munster)
- **UHL** — University Hospital Limerick (Midwest)
- **GUH** — Galway University Hospital (West)

## Realistic Missing Data

The generator introduces realistic patterns of missing data
to simulate real-world data quality issues:

```{r}
# Missing data rates for key variables
vars <- c("qblg9", "qblg3", "qblb1", "qblb4", "qblb9",
          "qbla9", "qbla4", "qble1", "qblf1", "qhd20")
missing_pct <- sapply(vars, function(v) {
  round(mean(is.na(df[[v]])) * 100, 1)
})
data.frame(Variable = vars, Missing_Pct = missing_pct)
```

## Using with Metric Functions

All metric compute functions accept the synthetic data directly:

```{r}
# Compute URR achievement by centre
urr <- compute_urr_achievement(df)
head(urr)
```

```{r}
# Compute haemoglobin achievement
hb <- compute_hb_achievement(df)
head(hb)
```

```{r}
# Compute missing data rates
missing <- compute_missing_data(df)
head(missing)
```

## Reproducibility

Setting the seed ensures identical datasets across runs:

```{r}
df1 <- generate_synthetic_data(seed = 123, n_total = 100)
df2 <- generate_synthetic_data(seed = 123, n_total = 100)
identical(df1, df2)
```

## Exporting for Power BI

The synthetic data can be exported as CSV for Power BI development:

```{r eval=FALSE}
readr::write_csv(df, "synthetic_audit_data.csv")
```

This is useful for the HSE IIS Power BI team to develop their
dashboard against realistic test data.
